Clean spectral data

### 2020-09-15 R/clean_spectral_data.Rmd
writing figures to files
noticed that the all_spectra file is smaller and has fewer samples than it should... guess I'll run the spectral-lib code again and try to see what's up

```{r}
# read_csv --> no factors, creates a tibble
all_spectra <- read_csv(here::here("analysis","data","raw_data",
                                   "spectral-lib","all_spectra.csv"))

# quick look at the values in each column
glimpse(all_spectra)
```

@@

Based on Max's Gist code: https://gist.github.com/mbjoseph/5c18781e508460e14f64193571b98b7d

@@

```{r, warning=FALSE, fig.height=16, fig.width=14}
all_spectra %>%
  ggplot(aes(wavelength_nm, ifelse(mask, NA, reflectance), group = spectraID,
             # color each facet plot based on taxonRank
             color = taxonRank)) +
  geom_path(alpha = .2) +
  labs(x = '\n Wavelength (nm)',
       y = 'Reflectance \n',
       color = "Taxon Rank") + 
  facet_wrap(~genusSpecies
             # wrap long names
             ,labeller = label_wrap_gen(20)) + 
  scale_color_manual(values = c("#e6550d","#1c9099", "#a8ddb5")) + 
  theme_minimal() + 
  # Adjust the text size and rotation angle 
  theme(text = element_text(size=11),
        axis.text.x = element_text(angle=90, hjust=1),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```

@@

```{r warning=FALSE, fig.width=12}
# plot spectra to see their starting wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra with different wavelengths: zoom in on beginning of spectra") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()
```

@@

```{r warning=FALSE, fig.width=12}
# plot spectra to see their ending wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra each with different wavelengths: zoom in on end of spectra") + 
  # zoom in on end of spectra
  lims(x = c(2455,2515), y = c(0, 0.3)) + 
  theme_bw()
```

@@

```{r, fig.width=12}
# plot interpolated spectra to see their starting wavelengths
interp_wl %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D11.CLBJ.00496_2018", 
                                 "NEON.PLA.D01.HARV.09041_2018",
                                 "NEON.PLA.D08.DELA.03903_2017")) %>% 
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra w interpolated wavelengths: Zoom in on beginning of spectra") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()
```

@@

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(taxonCounts, aes(y = reorder(genusSpecies, n), x = n)) + 
                          #reorder(taxonID, n), x = n)) + 
  geom_bar(stat = "identity", aes(fill = countGT)) + 
  labs(title = paste("Histogram of samples per species.",
                     speciesCountTotal, "total species;",
                     speciesCountGT, "species with greater than",
                     thresh, "samples"),
       y = "Genus Species",
       fill = paste("count greater than threshold", thresh)) + 
  theme_bw() + 
  scale_fill_manual(values = c("#d95f0e", "#7bccc4"))
```

>>>>

```{r}
# read_csv --> no factors, creates a tibble
all_spectra <- read_csv(here::here("analysis","data","raw_data",
                                   "spectral-lib","all_spectra.csv"))
```

@@

@@

```{r, warning=FALSE, fig.height=16, fig.width=14, eval=FALSE}
all_spectra %>%
  ggplot(aes(wavelength_nm, ifelse(mask, NA, reflectance), group = spectraID,
             # color each facet plot based on taxonRank
             color = taxonRank)) +
  geom_path(alpha = .2) +
  labs(x = '\n Wavelength (nm)',
       y = 'Reflectance \n',
       color = "Taxon Rank") + 
  facet_wrap(~genusSpecies
             # wrap long names
             ,labeller = label_wrap_gen(20)) + 
  scale_color_manual(values = c("#e6550d","#1c9099", "#a8ddb5")) + 
  theme_minimal() + 
  # Adjust the text size and rotation angle 
  theme(text = element_text(size=11),
        axis.text.x = element_text(angle=90, hjust=1),
        axis.title.y = element_text(size = 14),
        axis.title.x = element_text(size = 14),
        legend.text = element_text(size = 14),
        legend.title = element_text(size = 14))
```

@@

```{r warning=FALSE, fig.width=12}
# plot spectra to see their starting wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D10.RMNP.04210_2018", 
                                 "NEON.PLA.D03.DSNY.00878_2018",
                                 "NEON.PLA.D01.BART.03618_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra with different wavelengths: zoom in on beginning of spectra",
       x = "wavelength [nm]",
       color = "spectrum ID") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()

# save figure to file
ggsave(filename = here::here("analysis","figures", "wl_before_interp_380nm.pdf"), width = 9, height = 4)
```

@@

```{r warning=FALSE, fig.width=12}
# plot spectra to see their ending wavelengths
all_spectra %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D10.RMNP.04210_2018", 
                                 "NEON.PLA.D03.DSNY.00878_2018",
                                 "NEON.PLA.D01.BART.03618_2017")) %>%
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra each with different wavelengths: zoom in on end of spectra",
        x = "wavelength [nm]",
       color = "spectrum ID") + 
  # zoom in on end of spectra
  lims(x = c(2455,2515), y = c(0, 0.3)) + 
  theme_bw()

# save figure to file
ggsave(filename = here::here("analysis","figures", "wl_before_interp_2500nm.pdf"), width = 9, height = 4)
```

@@

```{r, fig.width=12}
# plot interpolated spectra to see their starting wavelengths
interp_wl %>% 
  dplyr::filter(spectraID %in% c(#"NEON.PLA.D20.PUUM.00954_2019",
                                 "NEON.PLA.D10.RMNP.04210_2018", 
                                 "NEON.PLA.D03.DSNY.00878_2018",
                                 "NEON.PLA.D01.BART.03618_2017")) %>% 
  ggplot(aes(x = wavelength_nm, y = reflectance, group = spectraID)) +
  geom_line(aes(color = spectraID), 
            alpha = 0.5, size = 1) +
  labs(title = "Spectra w interpolated wavelengths: Zoom in on beginning of spectra",
       x = "wavelength [nm]",
       color = "spectrum ID") + 
  # zoom in on beginning of spectra
  lims(x = c(380,400), y = c(0,0.1)) +
  theme_bw()

# save figure to file
ggsave(filename = here::here("analysis","figures", "wl_after_interp_380nm.pdf"), width = 9, height = 4)
```

@@

# plot histogram with count per genusSpecies. color bars based on whether
# # samples exceeds the threshold
ggplot(taxonCounts, aes(y = reorder(genusSpecies, n), x = n)) + 
                          #reorder(taxonID, n), x = n)) + 
  geom_bar(stat = "identity", aes(fill = countGT)) + 
  labs(title = paste("Histogram of samples per species.",
                     speciesCountTotal, "total species;",
                     speciesCountGT, "species with greater than",
                     thresh, "samples"),
       y = "Scientific Name (Genus + Species)",
       fill = paste("Count >", thresh)) + 
  theme_bw() + 
  scale_fill_manual(values = c("#d95f0e", "#7bccc4"))

# save figure to file
ggsave(filename = here::here("analysis","figures", "hist_samples_per_species.pdf"), width = 10, height = 12)
```
